Naturalness and Long-lived Beyond the Standard Model Particles

### 2017-07-18 tex/draft.tex
7/17 update

\section{The Standard Model}

>>>>

\section{The Standard Model}
    The standard model of particle physics (SM) decribes all known particles and forces (except gravity) and is experimentally verified to astonishing (*weak word) accuracy. The SM consists of 12 spin-1/2 particles, the fermions, that make up all observed matter, 5 spin-1 particles, the gauge bosons, that communicate the electromagnetic, weak, and strong fundamental forces, and 1 fundamental scalar, the Higgs boson, which will be described in detail below.

    The SM particles fit into multiplets of three symmetry groups, which makes the SM an SU(3)cxSU(2)LxSU(1)Y theory. These three symmetry groups correspond to color, the charge associated with the strong nuclear force; weak isopin, the charge associated with the left-handed chiral doublet something something; and weak hypercharge, the something something something, respectively.

    The Higgs potential (*insert equation) has a nonzero vacuum expectation value (VEV), which spontaneously breaks the SU(2)LxU(1)Y symmetry to a U(1)EM symmetry. Breaking the continuous symmetries creates 4 Nambu-Goldstone bosons, 3 of which are 'eaten' by the W+, W-, and Z bosons, causing the gauge bosons to acquire mass (*say in terms of DOF instead?). After parameterizing the Higgs field in terms of its VEV and propagating something something, the SM Lagrangian now contains mass terms for the quarks and leptons (*this is the first time I'm saying those words)(*Include eqs probably).

    Like all SM particles, the Higgs boson interacts with virtual particles in the vacuum through loop processes. The Higgs is unique, however, in that it is the only fundamental scalar in the theory. As a fundamental scalar, the Higgs sees corrections to its mass that depend quadratically on the energy scale at which the SM is replaced by a more complete theory (*eqs). The fermions and gauge bosons do not suffer this same fate because their masses are protected by the chiral and gauge symmetries, which are referred to as custodial symmetries in this context. The Higgs mass is found experimentally to be 125.1 GeV, even though it's sensitive to energy scales some 10\^16 times higher. This leads many (*weak) to describe the Higgs mass as 'unnatural', a concept that will be defined thoroughly below.
